Abstraction and Invariance for Algebraically Indexed Types

### 2012-07-06 paper/section-general-framework.tex
Definitions of contextual equivalence, semantic equivalence and the soundness property

@@

>>>>

We use our index-erasure semantics to define when a pair of programs
are contextually equivalent with respect to syntactically defined
applicative contexts.

\begin{definition}\label{defn:ctxt-equiv}
  Let $\Gamma_{\mathit{ops}}$ be a typing context with no free index
  variables that describes the types of the primitive operations, and
  let $\eta_{\mathit{ops}} \in \ctxtSem{\Gamma}$ be an interpretation
  of these primitive operations.

  Two programs $\Delta; \Gamma_{\mathit{ops}}, \Gamma \vdash M_1 : A$
  and $\Delta; \Gamma_{\mathit{ops}}, \Gamma \vdash M_2 : A$ are
  \emph{contextually equivalent} ($\Delta; \Gamma_{\mathit{ops}},
  \Gamma \vdash M_1 \stackrel{ctx}\approx M_2 : A$) if for all
  programs $-;\Gamma_{\mathit{ops}} \vdash T : (\forall \Delta. \Gamma
  \to A) \to \tyPrimNm{unit} + \tyPrimNm{unit}$, it is the case that
  $\tmSem{T\ (\Lambda \Delta. \lambda \Gamma. M_1)}\eta_{\mathit{ops}}
  = \tmSem{T\ (\Lambda \Delta. \lambda
    \Gamma. M_2)}\eta_{\mathit{ops}}$.
\end{definition}

@@

We use the relational interpretation to define the notion of semantic
equivalence, which is a sound approximation of contextual equivalence,
as a consequence of \thmref{thm:abstraction}.
\begin{definition}
  Let $\Gamma_{\mathit{ops}}$ and $\eta_{\mathit{ops}}$ be a context
  of primitive operations and its interpretation as in
  \defref{defn:ctxt-equiv}. Let $\relEnv{E}_{\mathit{ops}}$ be a
  family of sets of relational environments such that for all $\Delta$
  and $\rho \in \relEnv{E}_{\mathit{ops}}(\Delta)$,
  $(\eta_{\mathit{ops}}, \eta_{\mathit{ops}}) \in
  \rsem{\Gamma}{\relEnv{E}_{\mathit{ops}}}\rho$.

  Two programs $\Delta; \Gamma_{\mathit{ops}}, \Gamma \vdash M_1 : A$
  and $\Delta; \Gamma_{\mathit{ops}}, \Gamma \vdash M_2 : A$ are
  \emph{semantically equal} ($\Delta; \Gamma_{\mathit{ops}}, \Gamma
  \vdash M_1 \stackrel{sem}\sim M_2 : A$) if for all $\rho \in
  \relEnv{E}_{\mathit{ops}}$, and all $(\eta_1,\eta_2) \in
  \rsem{\Gamma}{\relEnv{E}_{\mathit{ops}}}\rho$, we have
  $(\tmSem{M_1}(\eta_{\mathit{ops}},
  \eta_1),\tmSem{M_2}(\eta_{\mathit{ops}}, \eta_2)) \in
  \rsem{A}{\relEnv{E}_{\mathit{ops}}}\rho$.
\end{definition}

\begin{theorem}\label{thm:soundness}
  $\Delta; \Gamma_{\mathit{ops}}, \Gamma \vdash M_1 \stackrel{sem}\sim M_2 : A$ implies
  $\Delta; \Gamma_{\mathit{ops}}, \Gamma \vdash M_1 \stackrel{ctx}\approx M_2 : A$
\end{theorem}
